---
title: "Choosing mtDNA screen hits for validation screen"
output: html_notebook
---



### Loading data and dependencies

```{r}
setwd("/Users/jtm/OneDrive - Johns Hopkins/li_lab/barseq/analysis/analysis_notebook")
getwd()
library(ggplot2)
library(ggrepel)
library(ggthemes)
library(ggpubr)
theme_set(theme_pubr())

load('outputs/de_results.rdata')
de_results <- de_results[de_results$condition =='m2',]

# Subset mtDNA data
de_results$sign <- sign(de_results$logFC)
de_results$rank_metric <- -log10(de_results$PValue)*de_results$sign
plot(de_results$logFC, de_results$rank_metric)
cor(de_results$logFC, de_results$rank_metric, method = 'pearson')
hist(-log10(de_results$PValue))
hist(de_results$PValue)
hist(de_results$rank_metric)
      
plot(de_results$logFC, -log10(de_results$PValue))
      
# Rank metric is signed negative log of the p-value
      
# Have some that are the negative signed? 
      
      
      
```

### Plot

```{r}
pval_cutoff <- 1e-9
forplot <- de_results
plot <- ggplot(data = forplot, aes()) +
      geom_point(aes(x=logFC, y=rank_metric, color='blue')) +
      geom_hline(yintercept=-log(pval_cutoff), color='red') + # Horizontal line for pval cutoff
      geom_hline(yintercept=log(pval_cutoff), color='red') + # Horizontal line for pval cutoff
      geom_hline(yintercept=300, color='red') + # Horizontal line for pval cutoff
      theme(plot.title = element_text(hjust=0.5)) +
      #theme(plot.subtitle  = element_text(hjust=0.5)) +
          theme(legend.position = "none") +
      xlab('log(Fold-chage)') +
      ylab('log(P-value)') +
      #xlim(-1,1) +
      ggtitle(label = 'Differential Barcode Frequencies \n Proliferation Lacking mtDNA') +
      geom_text_repel(
          fontface = 'bold',
          data=subset(forplot,
                  abs(rank_metric) > 150),
          aes(x=logFC, y=rank_metric,
          label=common),
          force = 20,
          show.legend = FALSE)
plot
```
# Subset by rank metric and plot results 

```{r}

# Define validation set as 45 lowest on rank, 147 highest on rank metric
de_results <- de_results[order(de_results$rank_metric,  decreasing = TRUE), ] 
validation_set <- rbind(head(de_results, 147), tail(de_results, 45))


# Sanity check plots 
rejected_set <- de_results[150:3300,] #Define a rejected set to get bounds on it. Approximate.
plot(1:100, runif(100,1,100))
plot(validation_set$logFC, validation_set$rank_metric, xlim=c(-5,5), ylim=c(-300, 300))
plot(rejected_set$logFC, rejected_set$rank_metric, xlim=c(-5,5), ylim=c(-300, 300))

pval_cutoff <- 1e-9
forplot <- validation_set
plot <- ggplot(data = forplot, aes()) +
      geom_point(aes(x=logFC, y=rank_metric, color='blue')) +
      geom_hline(yintercept=-log(pval_cutoff), color='red') + # Horizontal line for pval cutoff
      geom_hline(yintercept=log(pval_cutoff), color='red') + # Horizontal line for pval cutoff
      theme(plot.title = element_text(hjust=0.5)) +
      #theme(plot.subtitle  = element_text(hjust=0.5)) +
          theme(legend.position = "none") +
      xlab('log(Fold-chage)') +
      ylab('log(P-value)') +
      #xlim(-1,1) +
      ggtitle(label = 'Differential Barcode Frequencies \n Proliferation Lacking mtDNA') +
      geom_text_repel(
          fontface = 'bold',
          data=subset(forplot),
          aes(x=logFC, y=rank_metric,
          label=common),
          force = 20,
          show.legend = FALSE)
plot

# Jitter plot
forplot$rand <- runif(dim(forplot)[1],1,10)
plot <- ggplot(data = forplot, aes()) +
      geom_point(aes(x=rank_metric, y=rand, color='blue')) +
      #theme(plot.subtitle  = element_text(hjust=0.5)) +
          theme(legend.position = "none") +
      xlab('log(Fold-chage)') +
      ylab('log(P-value)') +
      #xlim(-1,1) +
      ggtitle(label = 'Differential Barcode Frequencies \n Proliferation Lacking mtDNA') +
      geom_text_repel(
          fontface = 'bold',
          data=subset(forplot),
          aes(x = rank_metric, y = rand,
          label=common),
          force = 20,
          show.legend = FALSE)
plot

```



I should do rank metric on the other datasets too 


# How to set up the plate 
```{r}
print('Available wells per Tecan plate, eliminating the perimeter and 1 WT well')
wells <- 96 - 12 - 12 - 6 - 6 - 1
wells

```
Let's do 1 plate of negative, 2 of positive
Shall i include more control wells per plate? 

No just do 50 lows and remainder highs and set plates for experiment later.

I need to make the control plasmid empty or finish sequencing this one...



### Output a list for pathway analysis and GSEA
```{r} 
# GSEA using BioGrid and bioconductor
gene_ranks <- de_results[,c(1,9)]


```


### Output a list for pathway analysis and picking clones (first 2 plates)
```{r}
# Select gene name and rank metric (signed -log10(PValue)
for_printer <- validation_set[,c(1,9)]
plate_map <- read.csv(file = 'inputs/barcode_key.csv', header = TRUE)
plate_map <- plate_map[, c(2, 3, 4)]
well_factors <- as.vector(read.csv(file = 'inputs/well_factors.csv', header = FALSE))
well_factors <- well_factors[,1]
plate_factors <- read.csv(file = 'inputs/plate_factors.csv')
plate_factors <- plate_factors[,1]
# Add plate positions

for_printer <- merge(x = for_printer, plate_map, by='gene')
# a$dup <- !a$gene %in% a$gene[duplicated(a$gene)] # Look for duplicates in the df
for_printer <- for_printer[-62,]
for_printer$well <- factor(for_printer$well, levels = well_factors)
for_printer$plate <- factor(for_printer$plate, levels = plate_factors)

for_printer <- for_printer[order(for_printer$well, decreasing = FALSE),] 
for_printer <- for_printer[order(for_printer$plate, decreasing = FALSE),] 

write.csv(for_printer, file = 'outputs/mtDNA_validation_set_picking_map.csv')
          
```









